Value-Suppressing Uncertainty Maps

### 2017-03-31 paper/template.tex
Typo pass

We describe the motivations behind the use of VSUMs, examples of their utility for decision-making under uncertainty, and assess VSUMs in a crowdsourced experiment. Our experimental results indicate that VSUMs promote better integration between uncertainty and data, promoting rapid but cautious decision-making.

@@

There are a number of techniques for communicating values and uncertainty in one-dimensional statistical graphics~\cite{olston2002visualizing, wickham201140, neyman1937outline}. Techniques such as error bars, confidence bands, and violin plots~\cite{hintze1998violin} use a separate position channel to encode uncertainty. The techniques in this paper also apply to 2D visualizations such as heatmaps. Today, there are no established standards of visualizing uncertainty in heatmaps.

To support our goal of making uncertainty information well-integrated with data, we focus on the techniques of \emph{juxtaposition} (where uncertainty is presented alongside the data), \textbf{superposition} (where additional glyphs are overlaid on top of the data), and \emph{explicit encoding} (where an additional visual channel is used to encode the uncertainty). Gleicher \ea~\cite{gleicher2011visual} explore these three strategies from the perspective of affording \emph{comparisons}; here, we examine them from the perspective of uncertainty visualization.

@@

\emph{Explicit encoding} directly represents uncertainty with a visual channel, simultaneously with data. VSUMs employ this strategy. In addition to overcoming the drawbacks of the other techniques mentioned in this section, we believe that explicit encoding represents the most general form of the solution of how to fuse uncertainty information with data. The most pressing design decisions for explicit encoding are 1) which visual variable to use to encode uncertainty and 2) how to design bivariate maps which afford the encoding of both value and uncertainty. We deal with these two questions in greater detail below.

@@

For visualizations such as choropleth maps, heatmaps, or treemaps, visual channels such as position and length are reserved for data variables other than value (such as geographic location or relative size). In these situations, data value is often encoded using color. Colors in such univariate quantitative color maps should be sufficiently far apart as to be perceptually distinguishable~\cite{ware1988color}, and vary in lightness as well as hue to afford an implicit ordering of value~\cite{borland2007rainbow,rogowitz2001blair}. Different choices of color maps can highlight different features of the data, and should be chosen with care based on the data types and important features to be visualized~\cite{rogowitz1996not}.

Given these considerations, and the perceptual integrality of color channels, the construction of \emph{bivariate} color maps is difficult. Viewers need to be able to independently assess the variables of interest, but the resulting bivariate map should also be expressive and easy to interpret. Bivariate maps with these constraints are often limited to relatively few categories (say, a 3x3 matrix as in \figref{fig:example}) \cite{robertson1986generation,trumbo1981theory}. In general, the quality of bivariate color maps is a multivariate measure involving consideration of not just the component color channels, but also the interpolation scheme and the color of the surround~\cite{bernard2015survey}. However, even simple bivariate maps can be difficult for a general audience to interpret \cite{wainer1980empirical}.

Dunn~\cite{dunn1989dynamic} attempts to circumvent these issues by dynamically allocating bins to the bivariate map. As the number of distinct, separable colors is limited, the dynamic approach allocates bins to regions of the highest utility. For instance, placing the dividing line between bins in areas aligned with a trend line allows the color to encode information about the sign of the residual with respect to each variable. This insight, that the bins in a bivariate map are a limited resource and so should be ``spent'' wisely, drives the creation of VSUMs, which allocate more bins in a bivariate map to regions with less uncertainty.

@@

Correll \ea~\cite{correll2015layercake,correll2011visualizing} use a pre-cursor of VSUMs in their LayerCake genomics visualization tool, where marks representing genomic data with increasing uncertainty are mapped to a smaller and smaller set of increasingly grey colors, creating the effect of uncertain values retreating into ``fog'' while highly certain values remain prominent. Other bivariate visualizations implicitly alias together uncertain values. For instance, if uncertainty is encoded by transparency, a maximally uncertain glyph may be entirely transparent, and so impossible to distinguish from any other maximally uncertain glyph. Other channels with a semiotic connection to uncertainty, such as saturation, value, blur, or size, also have deleterious effects on the disambiguation of colors and shapes. In both cases, the property of aliasing is \emph{ad hoc}, and places no guarantees on the discriminability of colors. The binning and degradation approach of VSUMs makes the choice to alias values explicit to both the designer and the viewer, and results in a bivariate mapping with known perceptual properties.

@@

How quickly the number of bins decreases is another parameter of VSUMs. If this degradation is too large, then we may only have a few discriminable levels of uncertainty. If this degradation is too small, then we can be overwhelmed with colors. For instance, if we only decrement by one bin per level of uncertainty, there would be \emph{1,485} different colors in a VSUM with a tolerance threshold of $1$ unit in CIELAB space. In the examples in this paper, we halve the number of bins at each increasing level of uncertainty. When the initial bins at the highly certain region is a power of $2$ (as in Fig. \ref{fig:example}), this imparts a tree structure to the resulting VSUM. As uncertainty increases, bins in the previous level map to one and only one ``parent'' in the next uncertainty level, resulting in fewer crossings of color boundaries among rows.

@@

Every pair of colors quantized color maps should be easily distinguishable from any other color in the color map. VSUMs have a lower resolution for uncertain values and thus can have higher resolution for values with high certainty. We created 2D uncertainty maps and VSUMs and used them to visualize flight delay data and mutations in viral genomes to show that VSUMs have higher discriminability where it matters.

The airline delay visualization shows average delays for different times of the day and days of the week. In this example, we use data about flight delays of US carriers from the Bureau of Transportation Statistics~\cite{bts} for January 2017. In addition to the average delay for each pair of time of the day and day of the week, we computed the standard error of the mean as the standard deviation divided by the squareroot of the number of samples.

@@

\figref{fig:airline2d} shows the heatmap visualization where the delay and uncertainty are encoded with a 2D uncertainty map. We used a minimum discriminability threshold of $18$ units in CIELAB space. We can see that the lowest delays are early in the day throughout the week. With the same threshold we can create a VSUM with much higher resolution for values with high certainty (\figref{fig:airlineVsum}). The VSUM also has more resolution for the uncertainty (4 different levels instead of 3 in the 2D heatmap). If the analyst is interested in delay patterns, VSUMs give them higher resolution.

@@

In \figref{fig:viral2d} and \figref{fig:viralVsum}, we visualized the variability in viral genomes across 25 populations of the simian immunodeficiency virus (SIV)~\cite{o2012conditional}. Next generation sequencing (NGS) allows biologists to collect and analyze large amounts of genomics data. However, these techniques often create data quality issues, as they must align large numbers of potentially ambiguous ``reads'' of base pairs. In the dataset here, it is important not only to identify hotspots in the viral genome (locations with high rates of mutation or variability), but to not be distracted by false positives. Extending the ``confidence fog'' metaphor used in viral genome browsers like LayerCake~\cite{correll2015layercake}, VSUMs enable biologists to explore interesting patterns while not being distracted by noise, as shown in \figref{fig:viralVsum}.

@@

\begin{enumerate}
	\item [T1] an \textbf{identification} task, where we gave participants charts with value and uncertainty information, and asked them to locate specific regions. E.g., ``click on the region of the chart with highest uncertainty \emph{and} lowest value.''
	\item [T2] a \textbf{prediction} task, similar to Battleship, where we gave participants charts with both \emph{forecast} and \emph{forecast uncertainty} information, and asked the participants to place tokens on the board in order to optimize expected value. E.g., ``place your $4$ ships on safe locations on the board.''
\end{enumerate}

@@

For each task, participants saw one of three types of thematic maps. Either two \textbf{juxtaposed} maps side-by-side, with separate encodings for value and uncertainty, a traditional \textbf{bivariate} map, where value and uncertainty are orthogonally encoded, or a \textbf{VSUM}. \figref{fig:conditions} shows examples of these factor levels. Other factors included the size of the map (either a 4x4 or 8x8 grid), and (for the prediction task), the question framing (attack or defense). All factors were presented within subjects. Secondary factors were evenly blocked across stimuli. Fatigue effects were significant in piloting, and so we limited the number of replicates: participants saw 18 stimuli (6 for each map type) for the identification task, and 12 (4 for each map type) for the prediction task.

In all cases, we generated color maps based on the requirement that no two colors in the resulting color scheme were closer than 18 units of Euclidean distance in CIELAB space. This threshold results in a $9$ color bivariate map, $15$ color VSUM for all color scales used in this study. Given the size of the marks used in our study, this threshold distance is sufficient for colors to be correctly distinguished in over 95\% of trials~\cite{stone2014engineering}. Other choices of threshold, and differing univariate scales, would result in different maps (see Fig.~\ref{fig:performance}).

@@

For the identification task, we gave participants bivariate representations of uncertainty, and asked them in a written prompt to identify a location meeting a given criteria, e.g., ``Click on the location with the greatest uncertainty.'' There were three potential criteria, mapping to the three vertices of a VSUM: 1) high value, low uncertainty, 2) low value, low uncertainty, and 3) high uncertainty. Each map contained four locations, one of each combination of maximum and minimum value and uncertainty. The correct answer was always one of these locations. Other locations in the grid were created pseudo-randomly, with less extreme ranges of uncertainty and probability. We measured performance both in terms of accuracy (did the participant select the correct point) as well as response time.

@@

For the prediction task, we gave participants the rules of a game like Battleship. Greis \ea~\cite{greis2016decision} employ these game-like experimental tasks to frame uncertainty information, which can be abstract or complex, to the general audience. In our task, the participant and a (fictional) adversary have to place tokens representing ships on a spatial grid. Without direct access to the location of the opponent's ships, the player then has to place tokens representing missile strikes on their opponent's grid. Players had to place all their tokens before continuing. If a player fires a missile at a square containing a ship, then that ship is sunk. The objective is the maximize the number of enemy ships hit, while minimizing the number of your own ships that are hit. In our task, participants were given a chart encoding predictions on the location of enemy ships (in the attacking case) or potential missile strikes (in the defending case). The \emph{value} component of the prediction was the likelihood of a item being present in a particular region. The \emph{uncertainty} component was the confidence in this prediction.

We included both a \textbf{gain} framing (given predictions of where your opponent's ships are, where should you launch your missiles?), and a \textbf{loss} framing (given predictions of where your opponent's missiles will fall, where should you place your ships?) to account for systematic differences in human judgments under uncertainty. Tversky \& Kahneman~\cite{tversky1985framing} illustrate that framings in terms of gains or losses produce reliably different outcomes. In general, the function of the value of gains is concave, but is convex for losses~\cite{kahneman1979prospect}. That is, avoiding a large loss is often more valuable than striving for a similarly large gain. To avoid potential (in this case, incorrect) transfer learning, the attacking and defending used different value color maps (Plasma and Cool) than the identification tasks, and from each other.  \figref{fig:taskTwoConditions} shows these differing color maps and framings.

@@

The ideal strategy from a value-maximizing standpoint would be to place tokens on areas with the highest expected value, ignoring the uncertainty information. However, as with Roulette and other similar games of chance, the variability in expected value is relevant when considering where to place bets~\cite{mlodinow2009drunkard}. Over the short term, a player may seek out high expected value but risky locations, or, if risk averse, accept lower expected value but lower variability locations. We therefore measured the distribution of both \emph{value} and \emph{uncertainty} of the tokens placed by the participants.

@@

We had several hypotheses, stemming from our believe that VSUMs promote better \emph{integration} between uncertainty and value information, and also encourage \emph{caution} by highlighting the ambiguity or untrustworthiness introduced by uncertain data. In particular:
\begin{enumerate}
	\item Participants would be \textbf{faster and more accurate} when completing the identification task using a VSUM as compared to the other map types. Juxtaposed maps, by adding an additional search task to the lookup task, would have especially poor performance compared to VSUMs.
	\item Participants would choose targets with \textbf{less uncertainty} in the prediction task when using a VSUM compared to the other maps types. This would result in a tradeoff where they would also choose targets with \textbf{less expected value} than the other map types. Again, juxtaposed maps, by increasing the difficulty of retrieving both variables simultaneously, would highlight this difference.
\end{enumerate}

@@

For the identification task, we were interested primarily in primary measures of performance, accuracy and response time. Our hypothesis was that VSUMs would have increased performance values compared to the other charts.

We performed a one-way ANOVA investigating the effect of chart type (juxtaposed, full bivariate, or VSUM) on accuracy. We found only a marginal effect ($F(2,859) = 2.4$, $p=0.09$), although VSUMs had the highest accuracy. \figref{fig:accuracy1} illustrates this result.

@@

We performed a one-way ANOVA of chart type on average value of guesses. We found no significant effect ($F(2,861)=0.27$, $p=0.8$), indicating that participants were generating similar expected values from their guesses across conditions. However, an ANOVA of chart type on average \emph{uncertainty} of guesses did find a significant main effect ($F(2,861)=61$, $p<0.001$). A post-hoc Tukey's HSD confirmed that participants had guesses with significantly less uncertainty with VSUMs than with both other chart types. Additionally, participants' guesses had significantly more uncertainty with juxtaposed charts than the other two. \figref{fig:uncertainty2} illustrates this result.

@@

Our results indicate that VSUMs promote avoidance of uncertain values, while not noticeably negatively impacting the expected value. Juxtaposed maps, by contrast, make the retrieval of uncertainty information significantly more difficult, and so it is possible that participants ignore or downweight it when making predictions. By making the consideration of uncertainty values integral to the design of the chart, VSUMs promote caution and consideration of uncertainty when making decisions.

@@

Bivariate maps have an inherent limitation in that visual channels are often difficult to attend to separately or orthogonally. Bivariate maps are also complex to interpret. Systems which simultaneously display large amounts of value and uncertainty data to general audiences (such as Pangloss~\cite{moritz2017trust}) avoid them for precisely these reasons, relying instead on juxtaposed maps. Our results indicate that juxtaposition produces additional problems, forcing users that wish to integrate uncertainty and data to locate an item of interest, and then re-locate that same item in a visually distant map, which may have few visual landmarks in common. This additional step before information fusion is reflected in longer response times, and riskier patterns of decision-making. Given the deleterious effects of latency on data analysis~\cite{liu2014effects}, designers ought to consider carefully the tradeoffs between expressiveness of visual encoding, and usability. VSUMs represent a middle ground between these two approaches, and have some of the benefits of both.

\subsection{Limitations and Future Work}
% When should you use these things?
In general, VSUMs excel in scenarios where we wish to make firm guarantees about the discriminability of different marks in a chart, but must also represent, in as much detail as possible, two, potentially orthogonal variables of interest. VSUMs have improved data resolution over na\"ive bivariate maps by assuming that uncertain values ought to have less visual impact that highly certain values. VSUMs therefore act both as a kind of filtering device as well as a bivariate representation. This assumption holds in scenarios like the virology dataset (\figref{fig:viralVsum}), where the expectation is that highly uncertain data is ``junk'' or should otherwise be downweighted in analysis. If the analyst has a different interpretation of uncertainty, or wishes to quickly and orthogonally analyses the distributions of uncertainty and value, other strategies, such as juxtaposed maps, may be more appropriate. VSUMs are designed for the \emph{integration} of value and uncertainty. Designers should take care in considering how when and how this integration is desirable.

@@

Often, uncertainty, data quality, or confidence are considered separate information from the data itself, relegated to tooltips or visually distant supplemental charts. We believe, in contrast, that uncertainty information ought to be directly integrated with charts of data itself. This integration introduces additional complexity in the design and presentation of data. Value Suppressing Uncertainty Maps represent one strategy for dealing with this complexity, by assigning marks properties in a way that supports the disambiguation of values in data where uncertainty is low, but suppresses these judgments when uncertainty is high.

>>>>

We describe the motivations behind the use of VSUMs, examples of their utility for decision-making under uncertainty, and assess VSUMs in a crowdsourced experiment. Our experimental results indicate that VSUMs create close integration between uncertainty and data, promoting rapid but cautious decision-making.

@@

There are a number of techniques for communicating values and uncertainty in one-dimensional statistical graphics~\cite{olston2002visualizing, wickham201140, neyman1937outline}. Techniques such as error bars, confidence bands, and violin plots~\cite{hintze1998violin} use a separate position channel to encode uncertainty. The techniques in this paper also apply to 2D visualizations such as heatmaps. To our knowledge, there are no established standards of visualizing uncertainty in thematic maps (but see MachEachren~\cite{maceachren1992visualizing} for a glossary of potential techniques).

To support our goal of making uncertainty information well-integrated with data, we focus on the techniques of \emph{juxtaposition} (where uncertainty is presented alongside the data), \emph{superposition} (where additional glyphs are overlaid on top of the data), and \emph{explicit encoding} (where an additional visual channel is used to encode the uncertainty). Gleicher \ea~\cite{gleicher2011visual} explore these three strategies from the perspective of affording \emph{comparisons}; here, we examine them from the perspective of uncertainty visualization.

@@

\emph{Explicit encoding} directly represents uncertainty with a visual channel, simultaneously with data. VSUMs employ this strategy. In addition to overcoming the drawbacks of the other techniques mentioned in this section, we believe that explicit encoding represents the most general form of the solution of how to fuse uncertainty information with data. The most pressing design decisions for explicit encoding are 1) which visual variables to use to encode uncertainty and 2) how to design bivariate maps which afford the encoding of both value and uncertainty. We deal with these two questions in greater detail below.

@@

For spatial visualizations like choropleth maps, heatmaps, and treemaps, visual channels such as position and length are reserved for data variables other than value (such as geographic location or relative size). In these situations, data value is often encoded using color. Colors in such univariate quantitative color maps should be sufficiently far apart as to be perceptually distinguishable~\cite{ware1988color}, and vary in lightness as well as hue to afford an implicit ordering of value~\cite{borland2007rainbow,rogowitz2001blair}. Different choices of color maps can highlight different features of the data, and should be chosen with care based on the data types and important features to be visualized~\cite{rogowitz1996not}.

Given these considerations, and the perceptual non-orthogonality of color channels, the construction of \emph{bivariate} color maps is difficult. Viewers need to be able to independently assess the variables of interest, but the resulting bivariate map should also be expressive and easy to interpret. Bivariate maps with these constraints are often limited to relatively few categories (say, a 3x3 matrix as in \figref{fig:example}) \cite{robertson1986generation,trumbo1981theory}. In general, the quality of bivariate color maps is a multivariate measure involving consideration of not just the component color channels, but also the interpolation scheme and the color of the surround~\cite{bernard2015survey}. However, even simple bivariate maps can be difficult for a general audience to interpret \cite{wainer1980empirical}.

Dunn~\cite{dunn1989dynamic} attempts to circumvent these issues by dynamically allocating bins to bivariate maps. As the number of distinct, separable colors is limited, the dynamic approach allocates bins to regions of the highest utility. For instance, placing the dividing line between bins in areas aligned with a trend line allows the color to encode information about the sign of the residual with respect to each variable. This insight, that the bins in a bivariate map are a limited resource and so should be ``spent'' wisely, drives the creation of VSUMs, which allocate more bins in a bivariate map to regions with less uncertainty.

@@

Correll \ea~\cite{correll2015layercake,correll2011visualizing} use a precursor of VSUMs in their LayerCake genomics visualization tool, where marks representing genomic data with increasing uncertainty are mapped to a smaller and smaller set of increasingly grey colors, creating the effect of uncertain values retreating into ``confidence fog'' while highly certain values remain prominent. Other bivariate visualizations implicitly alias together uncertain values. For instance, if uncertainty is encoded by transparency, a maximally uncertain glyph may be entirely transparent, and so impossible to distinguish from any other maximally uncertain glyph. Other channels with a semiotic connection to uncertainty, such as saturation, value, blur, or size, also have deleterious effects on the disambiguation of colors and shapes. In both cases, the property of aliasing is \emph{ad hoc}, and places no guarantees on the discriminability of colors. The binning and degradation approach of VSUMs makes the choice to alias values explicit to both the designer and the viewer, and results in a bivariate mapping with known perceptual properties.

@@

How quickly the number of bins decreases is another parameter of VSUMs. If this degradation is too fast, then we may only have a few discriminable levels of uncertainty. If this degradation is too slow, then we can be overwhelmed with colors. For instance, if we only decrement by one bin per level of uncertainty, there would be $1,485$ different colors in a VSUM with a tolerance threshold of $1$ unit in CIELAB space, far too much complexity for a discrete color mapping. In the examples in this paper, we halve the number of bins at each increasing level of uncertainty. When the number of initial bins at the highly certain region is a power of $2$ (as in Fig. \ref{fig:example}), this imparts a tree structure to the resulting VSUM. As uncertainty increases, bins in the previous level map to one and only one ``parent'' in the next uncertainty level, resulting in fewer crossings of color boundaries among rows.

@@

We present a set of examples showing how VSUMs can be applied to realworld datasets. The decision of VSUMs to allocate color bins assymmetrically amongst value and uncertainty regions affords greater discriminability in regions of the chart that warrant closer scrutiny, while discouraging exploration of regions with noisy or unreliable signals.

\subsubsection{Air Travel Delay}

@@

Using a public data set from the U.S. Bureau of Transportation Statistics~\cite{bts}, we created a heatmap showing average flight delay for different times of the day and days of the week for U.S. carriers in January 2017. We use standard error ($\frac{\sigma}{\sqrt{n}}$) as our measure of uncertainty. Starting from an minimum discriminability threshold of $18$ units in CIELAB space, we created two alternate heatmaps of this delay information. \figref{fig:airline2d} shows a traditional bivariate map, whereas \figref{fig:airlineVsum} shows a VSUM generated under the same constraints.

While both maps illustrate a similar temporal trend: flight delays are smaller towards the beginning of the day, increasing on average over time. The traditional bivariate map affords only a coarse examination of this trend: two visible ``blocks'' of color, early in the day, and later in the afternoon. The VSUM, in contrast, has sufficient color resolution to show a quasi-linear trend. The additional color resolution for uncertainty information in VSUMs also promotes caution when considering outliers: the small number of flights on Saturdays, or departing late at night, make standard error for these delays relatively high, and the connected values more suspect than the rest of the chart.

@@

Using a dataset from a group of virologists interested in the population dynamics of viral mutation~\cite{o2012conditional}, we visualized the variability in viral genomes across 25 populations of the simian immunodeficiency virus (SIV). Next generation sequencing (NGS) allows biologists to collect and analyze large amounts of genomics data. However, these techniques often create data quality issues, as they must align large numbers of potentially ambiguous ``reads'' of relatively small numbers of base pairs. In the dataset here, it is important not only to identify hotspots in the viral genome (locations with high rates of mutation or variability), but to not be distracted by false positives. The traditional bivariate map (\figref{fig:viral2d}) affords the analysis of only a few, large hot spots. aExtending the ``confidence fog'' metaphor used in viral genome browsers like LayerCake~\cite{correll2015layercake}, VSUMs (\figref{fig:viralVsum}) enable biologists to explore interesting patterns while not being distracted by noise. Smaller scale ``hot spots'' that all be mapped to the same magenta color in the traditional map are visible as bright orange lines. Pink hot spots towards the end of the genome (past reference nucleotide $9,000$) that could be plausibly considered potential regions of interest in the 2D map, are encoded as highly uncertain and of ambiguous value in the VSUMs. In fact, these regions have systematically low coverage; the lack of reliable data for these regions should discourage analysts from giving their apparently pattern equal consideration from the  stronger signals earlier in the genome.

@@

\begin{enumerate}
	\item an \textbf{identification} task, where we gave participants charts with value and uncertainty information, and asked them to locate specific regions. E.g., ``click on the region of the chart with highest uncertainty \emph{and} lowest value.''
	\item a \textbf{prediction} task, similar to Battleship, where we gave participants charts with both \emph{forecast} and \emph{forecast uncertainty} information, and asked the participants to place tokens on the board in order to optimize expected value. E.g., ``place your $4$ ships on safe locations on the board.''
\end{enumerate}

@@

For each task, participants saw one of three types of thematic maps. Either two \textbf{juxtaposed} maps side-by-side, with separate encodings for value and uncertainty, a traditional \textbf{2D bivariate} map, where value and uncertainty are orthogonally encoded, or a \textbf{VSUM}. \figref{fig:conditions} shows examples of these factor levels. Other factors included the size of the map (either a 4x4 or 8x8 grid), and (for the prediction task), the question framing (attack or defense). All factors were presented within subjects. Secondary factors were evenly blocked across stimuli. Participants reported fatigue effects in piloting, and so we limited the number of replicates in this study: participants saw 18 stimuli (6 for each map type) for the identification task, and 12 (4 for each map type) for the prediction task.

In all cases, we generated color maps based on the requirement that no two colors in the resulting color scheme were closer than 18 units of Euclidean distance in CIELAB space. This threshold results in a $9$ color bivariate map, $15$ color VSUM for all color scales used in this study. Given the size of the marks used in our study, this threshold distance is sufficient for a predicted discriminability rate of over 95\%~\cite{stone2014engineering}. Other choices of threshold, and differing univariate scales, would result in different maps (see Fig.~\ref{fig:performance}).

@@

For the identification task, we gave participants bivariate representations of uncertainty, and asked to identify a location meeting a given criteria, e.g., ``Click on the location with the greatest uncertainty.'' There were three potential criteria, mapping to the three vertices of a VSUM: 1) high value, low uncertainty, 2) low value, low uncertainty, and 3) high uncertainty. Each map contained four locations, one of each combination of maximum and minimum value and uncertainty. The correct answer was always one of these locations. Other locations in the grid were created pseudo-randomly, with less extreme ranges of uncertainty and probability. We measured performance both in terms of accuracy (did the participant select the correct point) as well as response time.

@@

For the prediction task, we gave participants the rules of a game like Battleship. Greis \ea~\cite{greis2016decision} employ these game-like experimental tasks to assess how different visual designs communicate uncertainty information, which can be abstract or complex, to the general audience. In our task, the participant and a (fictional) adversary have to place tokens representing ships on a spatial grid. Without direct access to the location of the opponent's ships, the player then has to place tokens representing missile strikes on their opponent's grid. Players have to place all their tokens before continuing. If a player fires a missile at a square containing a ship, then that ship is sunk. The objective is the maximize the number of enemy ships hit, while minimizing the number of your own ships that are hit. In our task, participants were given a chart encoding predictions of the location of enemy ships (in the attacking case) or potential missile strikes (in the defending case). The \emph{value} component of the prediction was the likelihood of a item being present in a particular region. The \emph{uncertainty} component was the confidence in this prediction.

We included both a \textbf{gain} framing (the attacker's task: given predictions of where your opponent's ships are, where should you launch your missiles?), and a \textbf{loss} framing (the defender's task: given predictions of where your opponent's missiles will fall, where should you place your ships?) to account for systematic differences in human judgments under uncertainty. Tversky \& Kahneman~\cite{tversky1985framing} illustrate that framings in terms of gains or losses produce reliably different outcomes. In general, the function of the value of gains is concave, but is convex for losses~\cite{kahneman1979prospect}. That is, avoiding a large loss is often more valuable than striving for a similarly large gain. To avoid potential (in this case, incorrect) transfer learning, the attacking and defending used different value color maps (Plasma and Cool) than in the identification task, and from the opposite framing in this task. \figref{fig:taskTwoConditions} shows these differing color maps and framings.

@@

The ideal strategy from a value-maximizing standpoint would be to place tokens on areas with the highest expected value, ignoring the uncertainty information. However, as with roulette and other similar games of chance, the variability in expected value is relevant when considering where to place bets~\cite{mlodinow2009drunkard}. Over the short term, a player may seek out high expected value but risky locations, or, if risk averse, accept lower expected value but lower variability locations. We therefore measured the distribution of both \emph{value} and \emph{uncertainty} of the tokens placed by the participants.

@@

Across both the identification and prediction tasks, we had several hypotheses, stemming from our belief that VSUMs promote better \emph{integration} between uncertainty and value information, and also encourage \emph{caution} by highlighting the ambiguity or untrustworthiness introduced by uncertain data. In particular:
\begin{enumerate}
	\item Participants would be \textbf{faster and more accurate} when completing the identification task using a VSUM as compared to the other map types. Juxtaposed maps, by adding an additional search task to the lookup task, would have especially poor performance compared to VSUMs.
	\item Participants would choose targets with \textbf{less uncertainty} in the prediction task when using a VSUM compared to the other maps types. This would result in a tradeoff where they would also choose targets with \textbf{less expected value} than the other map types. Again, juxtaposed maps, by increasing the difficulty of retrieving both variables simultaneously, would highlight this difference.
\end{enumerate}

@@

For the identification task, our primary measures of performance were accuracy and response time. Our hypothesis was that VSUMs would have increased performance values compared to the other charts.

We performed a one-way ANOVA investigating the effect of chart type (juxtaposed, full bivariate, or VSUM) on accuracy. We found only a marginal effect ($F(2,859) = 2.4$, $p=0.09$), although VSUMs had the highest accuracy among the three. \figref{fig:accuracy1} illustrates this result.

@@

We performed a one-way ANOVA of chart type on average value of guesses. We found no significant effect ($F(2,861)=0.27$, $p=0.8$), indicating that participants were generating similar expected values from their guesses across conditions.

However, an ANOVA of chart type on average \emph{uncertainty} of guesses did find a significant main effect ($F(2,861)=61$, $p<0.001$). A post-hoc Tukey's HSD confirmed that participants had guesses with significantly less uncertainty with VSUMs than with both other chart types. Additionally, participants' guesses on juxtaposed charts were significantly riskier than with the other chart types. \figref{fig:uncertainty2} illustrates this result.

@@

While our results only partially support our results, in that we did not observe statistically different patterns of \emph{expected value}, our did indicate that VSUMs produce different patterns of \emph{acceptable uncertainty}. That is, VSUMs promote avoidance of uncertain values, while not noticeably negatively impacting the expected value. Juxtaposed maps, by contrast, make the retrieval of uncertainty information significantly more difficult, and so it is possible that participants ignore or downweight it when making predictions, resulting in riskier predictions overall. By making the consideration of uncertainty values integral to the design of the chart, VSUMs promote caution when making decisions.

@@

Bivariate maps have an inherent limitation in that visual channels are often difficult to attend to separately or orthogonally. Bivariate maps are also complex to interpret. Systems which simultaneously display large amounts of value and uncertainty data to general audiences (such as Pangloss~\cite{moritz2017trust}) avoid them for precisely these reasons, relying instead on juxtaposed maps. Our results indicate that juxtaposition produces additional problems, forcing users that wish to integrate uncertainty and data to locate an item of interest, and then re-locate that same item in a visually distant map, which may have few visual landmarks in common. This additional step before information fusion is reflected in longer response times, and patterns of decision-making that are less mindful of risk. Given the deleterious effects of latency on data analysis~\cite{liu2014effects}, designers ought to carefully consider the tradeoffs between expressiveness of visual encoding, and usability. VSUMs represent a middle ground between these two approaches, and have some of the benefits of both.

\subsection{Limitations and Future Work}
% When should you use these things?
In general, VSUMs excel in scenarios where we wish to make firm guarantees about the discriminability of different marks in a chart, but must also represent, in as much detail as possible, two, potentially orthogonal variables of interest. VSUMs have improved data resolution over na\"ive bivariate maps by assuming that uncertain values ought to have less visual impact that highly certain values. VSUMs therefore act both as a kind of filtering device as well as a bivariate representation. This assumption holds in scenarios like the virology dataset (\figref{fig:viralVsum}), where the expectation is that highly uncertain data is unreliable or should otherwise be downweighted in analysis. If the analyst has a different interpretation of uncertainty, or wishes to quickly and orthogonally analyze the distributions of uncertainty and value, other strategies, such as juxtaposed maps, may be more appropriate. VSUMs are designed for the \emph{integration} of value and uncertainty. Designers should take care in considering how when and how this integration is desirable.

@@

Often, uncertainty, data quality, or confidence are considered separate information from the data itself, relegated to tooltips or visually distant supplemental charts. We believe, in contrast, that uncertainty information ought to be directly integrated with charts of data itself. This integration introduces additional complexity in the design and presentation of data. Value Suppressing Uncertainty Maps represent one strategy for dealing with this complexity, by assigning marks properties in a way that supports the disambiguation of values in data where uncertainty is low, but suppresses these judgments when uncertainty is high. This decision of how to allocate visual variables promotes patterns of decision-making that make responsible use of uncertainty information, discouraging comparison of values in unreliable regions of the data, and promoting comparison in regions of high certainty.
